---
title: "Salesforce Pipe Analysis"
author: "Douglas Martins"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyr)
require(dplyr)
require(ggplot2)
require(readr)
require(readxl)
require(stats)
require(yarrr)
require(lubridate)
options(dplyr.summarise.inform = FALSE)
```

## Introduction

The goal of this project is to identify attributes of contact data that correlate with over performance and under-performance in terms of funnel conversion

Funnel Stages: Inquiry → MQL → SAL → SQL → SQO → Won

Notes on nuance in the data

For all Campaigns except "US_FY17_HDC_SignUp", here are the definitions:

-   Inquiry - Exclude these as these contacts will not be worked by sellers\^1

-   MQL - Marketing qualified leads that are routed to sellers to be worked

-   SAL - Meeting is booked with the contact and an opportunity is created

-   SQL - Opportunity moved to at least stage 2 "Qualified"\^2

-   SQO - Opportunity moved beyond stage 2

-   Won - Either opportunity was closed won with Opportunity Amount \> \$0 OR opportunity was closed lost but closed lost reason = Signed Up for Access\^3

"US_FY17_HDC_SignUp" Campaign needs to be analyzed by itself as there is an automation in Salesforce that automatically moves contacts who confirm their emails in the customer's product from the MQL to SQL stage.

For these contacts, we need to look at SQO and Win conversion which have the same definitions as the other campaigns listed above.

Some Contacts are marked as disqualified. These contacts where Person or Contact Status = Disqualified should be removed from the analysis.

```{r input_data, include=FALSE, cache=TRUE}
contact_data <- readxl::read_xlsx(path = './data/contactattributescasestudy.xlsx',sheet = 1) %>% 
        #Here (3) is assumed to be correct since instructions were not clear
        mutate(lead_won=if_else(`Total Opportunity Amount`>0 | !is.na(`Sales Prospect / Sign Up`)
                                       ,1,0,missing=0)) %>%
        filter(`Person Status`!='Disqualified' | `Contact Status` != 'Disqualified')
account_data <- readxl::read_xlsx(path = './data/Data - Mapping.xlsx',sheet = 'FINAL PARENT')
win_percent <- round(mean(contact_data$lead_won),2)*100
```

Average win percentage is `r win_percent`%.

## Filtering Data

```{r data filtering 1, echo=FALSE}
#Trying a few percentages of missing records to filter columns
contact_data_dot50 <- ncol(contact_data[, which(colMeans(!is.na(contact_data)) > 0.5)])
contact_data_dot70 <- ncol(contact_data[, which(colMeans(!is.na(contact_data)) > 0.7)])
contact_data_dot90 <- ncol(contact_data[, which(colMeans(!is.na(contact_data)) > 0.9)])
```

In order to optimize time working on columns that have relevant information filtering columns that have a threshold of fill rate higher than an arbitrary percentage. The amounts were for 50%, 70% and 90%, respectively, 185, 72 and 60 attributes. The selected threshold is 90%.

```{r data filtering 2, echo=FALSE}
#Selecting a few columns from dataset
fil_cols <- unique(c(names(contact_data)[which(colMeans(!is.na(contact_data)) > 0.9)]))
contact_data_dot90 <- contact_data[,fil_cols]

#Checking uniques for each column
column_uniques_count <- data.frame(col=colnames(contact_data_dot90),uniques=apply(contact_data_dot90, MARGIN = 2, function(x){length(unique(x))}))
knitr::kable(column_uniques_count)

#Separating analysis columns
cat_cols <- c("Acquisition Program"
        ,"Campaign Type"
        ,"City"
        ,"IT Org Chart"
        ,"Sales Owner Email Address"
        ,"Person Source"
        ,"Lead Source Description"
        ,"MQL Day of Week"
        ,"Most Recent Campaign Member Status"
        ,"Original Source Type"
        ,"Person Time Zone"
        ,"Registration Source Info"
        ,"Registration Source Type"
        ,"SFDC Campaigns"
        ,"SFDC Type"
        ,"Sourced from Data.com"
        ,"State"
        ,"Usertype"
        ,"Job Level"
        ,"Region"
        ,'Original UTM Campaign'
        ,'Original UTM Content'
        ,'Original UTM Description'
        ,'Original UTM Medium'
        ,'Original UTM Source'
        ,'Original UTM Term'
        ,'UTM Campaign'
        ,'UTM Content (A)'
        ,'UTM Description'
        ,'UTM Medium'
        ,'UTM Medium (A)'
        ,'UTM Source'
        ,'UTM Term'
        ,'UTM Term (C)'
        ,'UTMCampaign'
        ,'UTMSource')
#Selecting analysis numeric cols
numeric_cols <- c("Admin Update Counter"
        ,"Behavior Score"
        ,"Job Function Score"
        ,"Job Level Score"
        ,"Person Score"
        ,"Number of Opportunities"
        ,"Response Score"
        ,"Response count")

contact_data <- contact_data[,c('lead_won','Marketing Funnel State','MQL Date',numeric_cols,cat_cols)]
```

With this in mind a few columns have been selected for a correlation analysis in the next section.

## Correlation Analysis

The analysis in this section will be split for each stage of the opportunity pipeline, aiming to find which attribute has highest correlation winning the lead. For numerical attributes the method selected to draw a comparison between itself and the success metric is ANOVA, as for categorical attributes, chi-squared test.

### MQL Stage

```{r}
mql_contact_data <- subset(contact_data,`Marketing Funnel State`=='MQL')
```

```{r}
mql_numeric <- sapply(numeric_cols, function(x,cdata=mql_contact_data){
        t_test <- t.test(formula = get(x) ~ factor(lead_won),data = cdata, paired=FALSE)
        x.aov <- aov(formula = get(x) ~ factor(lead_won), data = cdata)
        x.anova <- anova(x.aov)
        list(t_test$p.value, x.anova[1,5])
})
df_num <- as.data.frame(t(data.frame(mql_numeric))) %>% arrange(V1) %>% filter(V1<0.05)
colnames(df_num) <- c('t.test_pvalue','anova_p_value')
knitr::kable(df_num)
```

```{r}
mql_categorical <- sapply(cat_cols, function(x,cdata=mql_contact_data){
        cdata <- na.omit(cdata[,c(x,'lead_won')])
        if(nrow(unique(cdata[x]))>1 & length(unique(cdata$lead_won))>1){
                chisq_test <- chisq.test(cdata[x],cdata$lead_won, simulate.p.value = TRUE)
                list(chisq_test[3]$p.value)
        } else 1
})
df_cat <- as.data.frame(t(data.frame(mql_categorical))) %>% arrange(V1) %>% filter(V1<0.05)
colnames(df_cat) <- c('chisq.test_pvalue')
knitr::kable(df_cat)
```

### SQO Stage

```{r}
sqo_contact_data <- subset(contact_data,`Marketing Funnel State`=='SQO')
```

```{r}
sqo_numeric <- sapply(numeric_cols, function(x,cdata=mql_contact_data){
        t_test <- t.test(formula = get(x) ~ factor(lead_won),data = cdata, paired=FALSE)
        x.aov <- aov(formula = get(x) ~ factor(lead_won), data = cdata)
        x.anova <- anova(x.aov)
        list(t_test$p.value, x.anova[1,5])
})
df_num <- as.data.frame(t(data.frame(sqo_numeric))) %>% arrange(V1) %>% filter(V1<0.05)
colnames(df_num) <- c('t.test_pvalue','anova_p_value')
knitr::kable(df_num)
```

```{r}
sqo_categorical <- sapply(cat_cols, function(x,cdata=sqo_contact_data){
        cdata <- na.omit(cdata[,c(x,'lead_won')])
        if(nrow(unique(cdata[x]))>1 & length(unique(cdata$lead_won))){
                chisq_test <- chisq.test(cdata[,x],cdata$lead_won, simulate.p.value = TRUE)
                list(chisq_test[3]$p.value)
        } else 1
})
df_cat <- as.data.frame(t(data.frame(sqo_categorical))) %>% arrange(V1) %>% filter(V1<0.05)
colnames(df_cat) <- c('chisq.test_pvalue')
knitr::kable(df_cat)
```

In this step we noticed that in the MQL stage all numeric variables have a general correlation to the result (5%), but as for categorical, none of the UTM, Job, City and State parameters are statistically relevant. 

In the SQO step there is a notable appearence of the UTMSource variable as relevant.

As for specific analysis for each campaign, its possible to filter data and redo the analysis to see changes in the correlated variables.

## UTMSource conversion changes

```{r}
df <- contact_data %>% filter(!is.na(UTMSource)) %>%
        mutate(quarter=floor_date(ymd(`MQL Date`),'quarter')) %>% 
        filter(year(quarter)==2022) %>%
        group_by(UTMSource,quarter) %>% 
        summarise(.by = ,lead_won_rate=mean(lead_won))
p <- ggplot(df, aes(x=quarter, y=lead_won_rate, group=UTMSource, color=UTMSource)) +
  geom_line(linewidth=1.5) + geom_point() +
  xlab("")
p
```

Looking at the trends it seems that `myPresences` has a notable increase in conversion rate


